Energy and Momentum of a Density-Driven Overflow in the Samoan Passage

### 2023-02-04 reviews/response_to_reviewers.tex

\documentclass[11pt]{article}
\usepackage{graphicx}
% \usepackage[dvipsnames]{xcolor}
% \usepackage{palatino}
% \usepackage{mathpazo}
\usepackage{kpfonts}
\usepackage[top=1in, left=1in, bottom=1.1in, right=1in]{geometry}

% \pagestyle{plain}
% \topmargin 0mm
% \headsep 0mm
% \textwidth 6.5in
% \headheight 10mm
% \footskip 10mm
% \textheight 9in
% \oddsidemargin 0in
% \evensidemargin 0in

\newcommand{\dgr}{$^{\circ}$}

\begin{document}

\pagestyle{empty}

\begin{center}
{\large  SCRIPPS INSTITUTION OF OCEANOGRAPHY\\
UNIVERSITY OF CALIFORNIA, SAN DIEGO}\\
\vspace{0.1in}
{\small 9500 Gilman Drive \#0213\\
La Jolla, CA 92093}
\end{center}

\begin{flushright}
phone: 858.822.6809\\
email:~gvoet@ucsd.edu\\
\vspace{0.2in}
\today
\end{flushright}

\begin{flushleft}
% Addressee
Sylvia Cole, Editor\\
Journal of Physical Oceanography\\
American Meteorological Society\\
Boston, MA\\

\end{flushleft}

\vspace{.2in}

\noindent Dear Dr.\ Cole,

Please consider the enclosed manuscript, which was originally submitted as JPO-D-22-0220, ``Energy and Momentum of a Density-Driven Overflow in the Samoan Passage'' for consideration in \it Journal of Physical Oceanography\rm.  We have addressed all of the reviewer's comments, as seen in the following.

\begin{flushleft}
Sincerely,

\includegraphics[height=0.5in]{"/Users/gunnar/Documents/private/sig"}

%\vspace{.5in}
Gunnar Voet

\clearpage
In the following, we quote the reviewer's comments in italics, followed by our responses in bold.\\
\vspace{0.2in}

\centerline{\bf Responses to Reviewer 1}
\vspace{0.05in}

\it
This paper sets out the energetics of a dense water overflow at a key bottleneck in the northwards spreading of bottom water leading to identifying the key dynamical balances driving the flow and turbulent mixing. This is achieved using a blended approach of observations and numerical modelling allowing robust conclusions to be draw accounting for the limitations of each approach. The paper is very well written with a comprehensive description of the methodology and its limitations. Overall, I recommend this paper is returned for minor revisions to allow the authors to consider some minor comments (see below).

\vspace{0.05in}
\bf
We appreciate you taking the time to review this manuscript and would like to thank you for your kind words and helpful comments.

\begin{itemize}
\item \it
Line 102: change ``high resolution'' to ``high horizontal resolution'' to provide more clarity

\bf
Changed as suggested.

\item \it
Line 108 -- in the description of the sections there is no mention of 4a might be an oversight or intentional

\bf
We added the following sentence to highlight section 4a: ``Flow steadiness is discussed as a precursor to the following analysis in section~4a.''

\item \it
Figure 2 -- The vertical velocities in panel a are hard to interpret to this size. Perhaps a separate panel would make this clearer.

\bf
response

\item \it
Figure 4 -- Panel b is described as northwards velocity. I found this a little confusing as, if I understood correctly, the topography used for the model does not run directly North-South. Perhaps change the wording to across sill or horizontal velocity (as the model is 2D).

\bf
We changed the description for panel b to ``horizontal velocity'' as suggested.

\item \it
Equation 4 -- I believe ``$du$'' should be ``$\delta u$''.

\bf
Corrected. Thank you for pointing this out.

\item \it
Line 563 -- ``Fluxes within the dense layer'' I found it a little unclear here what fluxes you are referring to. I think it is the vertical pressure work however it could be the internal wave fluxes (both listed on line 559).

\bf
You are right, this should read ``Pressure work within the dense layer'' and is now corrected.

\item \it
Figure 13b -- You should say what offset has been used for the plot here so the reader can judge if the model and observations are showing similar magnitude anomalies If they are significantly different in magnitude a few words to explain why that is the case would be helpful.

\bf
Thank you for pointing this out. There is indeed a considerable offset of about 200\,N\,m$^{-2}$ between model and observations. This is caused by the way we initialize the model: It starts out with the observed density field but then drains some of the initial stratification associated with potential energy to accelerate the flow. We restore density at the domain boundaries, but over the sill the density structure has deviated somewhat from the initial values model hour~100 when we start the model analysis. This difference between model and observations is visible, for example, in Fig.~9a-c where the observations show higher potential energy than the model. This difference is now reflected in the caption.

\item \it
Line 948 -- The text here refers to the other choices of window size being shown in figure 11b however they are not. These should: be added to figure 11, given a new appendix figure, or the link to a figure removed.

\bf
Thank you for spotting this, the figure reference is a remnant of an earlier version of this manuscript. We removed the reference to the figure.

\item \it
Line 966 to 969 -- The text here states that the effect of diffusive fluxes is small. It would be helpful to have a quick scaling provided to support this statement as it is not (at least to me) obvious a priori given that in these deep ocean settings buoyancy flux can be an important sink of turbulent kinetic energy (e.g., Ijichi et al GRL 2020, Spingys et al JPO 2021).

\bf
response

\end{itemize}

%------------------------------------------

\clearpage

\centerline{\bf Responses to Reviewer 2}
\vspace{0.05in}

\it 
This paper examines detailed observations and a semi-idealized 2D (non-hydrostatic) model of a density-driven overflow. The observations are impressive (i.e., they include non-trivial estimates of TKE dissipation and vertical velocities) and the authors have carefully though about how to construct energy and momentum budgets given the strengths and limitations of the observations. The modeling results do not match the observations precisely, but provide a welcomed sanity check. The paper quantifies many useful terms in the energy and momentum budgets, which will allow quantitative comparisons with future observations, numerical simulations, and parameterizations. The analyses here have significant caveats and uncertainties, but all of these issues are well documented. The manuscript is well written and references the relevant literature. Many/most references are discussed in detail and linked to the central methods and results of the study (as opposed to a less-helpful pro-forma list of references in the introduction). This work will of be of interest the readers of JPO and I recommend the paper be published with minimal edits.
\vspace{0.05in}

\bf
Thank you very much for taking the time to review the manuscript and for your thorough comments. We are glad you found it interesting to read.

\begin{itemize}

\item \it
Fig. 9 Change the color bar on the 3rd and 4th row to match other fluxes? The purple color makes it look like the fluxes are going to the left, when compared with the rows below them.

\bf
Response
\end{itemize}

\clearpage

\end{flushleft}

\end{document}

### 2023-02-06 doc/northern_sill_energetics.tex

\abstract{
The energy and momentum balance of an abyssal overflow across a major sill in the Samoan Passage is estimated from two highly resolved towed sections, set 16 months apart, and results from a two-dimensional numerical simulation.
Driven by the density anomaly and associated pressure drop across the sill, the flow is relatively steady as apparent in the repeat sections presented in this study and in moored observations at several points along the flow presented in a previous paper.
The system gains energy from divergence of horizontal pressure work $\mathcal{O}(5)$\,kW\,m$^{-1}$ and flux of available potential energy $\mathcal{O}(2)$\,kW\,m$^{-1}$.
Approximately half of these gains are transferred into kinetic energy while the other half is lost to turbulent dissipation, bottom drag, and divergence in vertical pressure work.
Small-scale internal waves emanating downstream of the sill within the overflow layer radiate $\mathcal{O}(1)$\,kW\,m$^{-1}$ upwards but dissipate most of their energy within the dense overflow layer and at its upper interface.
The strongly sheared and highly stratified upper overflow interface thus acts as a critical layer and inhibits any appreciable upward radiation of energy via topographically generated internal lee waves.
Form drag of $\mathcal{O}(2)$\,N\,m$^{-2}$, estimated from the pressure drop across the sill, appears to be generally consistent with energy lost to dissipation and internal wave fluxes.
The topographic drag removes momentum from the mean flow, slowing it down and feeding a counter current aloft.
The processes discussed in this study combine to convert about one third of the energy released from the cross-sill density difference into turbulent mixing within the overflow and at its upper interface.
The observed and modeled vertical momentum flux divergence sustains gradients in shear and stratification, thereby maintaining an efficient route for abyssal water mass transformation downstream of this Samoan Passage sill.
}

@@

In this study we estimate the energy and momentum budget of flow across a major sill in the Samoan Passage using high resolution, towed, observations.
Results from a two-dimensional model are used to corroborate the analysis.
In the following, we give a short overview of the abyssal flow through the Samoan Passage and one of its major overflows (section~\ref{sec:ExperimentalDetails}\ref{sec:StudyRegion}), present towed observations of this overflow (section~\ref{sec:ExperimentalDetails}\ref{sec:Observations}),  and outline the setup of a two-dimensional numerical model simulating the dense overflow (section~\ref{sec:ExperimentalDetails}\ref{sec:Model}) to help interpret the observations.
After discussing the energy of the flow in terms of the Bernoulli equation (section~\ref{sec:Energetics}\ref{sec:Bernoulli}), a baroclinic energy equation is introduced (section~\ref{sec:Energetics}\ref{sec:EnergyEquation}).
Both frameworks are applied to the observed and modeled overflow for energy budgets in section~\ref{sec:Results}\ref{ssec:EnergyBudgetResults}.
Form drag is calculated and evaluated against the energy budgets (\ref{sec:Results}\ref{ssec:FormDrag}).
Upward momentum flux estimates are presented in section~\ref{sec:Results}\ref{ssec:MomentumFluxes}.
The results are discussed and compared to observations of other high drag flows in section~\ref{sec:Discussion}.

@@

\begin{figure*}
\centerline{\includegraphics[width=39pc]{model_snapshot.pdf}}
\caption{Model snapshot after 100 hours spinup time. Potential temperature is contoured in all four panels in 0.05$^{\circ}$C intervals starting at 0.7$^{\circ}$C. Thicker contours show the 0.8 and 0.9$^{\circ}$C isotherms.
(a) Potential temperature.
(b) Northward velocity.
(c) Vertical velocity.
(d) Turbulent dissipation $\varepsilon$ based on a parameterization acting on vertical density instabilities \citep[KL10,][]{klymaklegg10}.
}
\label{fig:ModelSnapshot}
\end{figure*}

@@

\subsection{Bernoulli flux}\label{sec:Bernoulli}
Treating the overflow as a single layer flow with the waters above at rest, we start out with the steady shallow water equations in one dimension:
\begin{equation}
    v \frac{\partial v}{\partial y} + g^\prime \frac{\partial}{\partial y} (\delta + h) = 0\ ,
    \label{eq:ShallowWater}
\end{equation}
where $\delta$ is the thickness of the layer, $v$ is the horizontal layer velocity along coordinate $y$, $h$ is the elevation of the topography, and $g^\prime=g \Delta \rho / \rho$ expresses the density difference $\Delta \rho$ between the bottom layer and waters aloft.
Neglecting entrainment leads to constant volume transport $Q$ of the dense bottom layer:
\begin{equation}
\frac{\partial (v \delta)}{\partial y} = \frac{\partial Q}{\partial y} = 0\ .
\end{equation}
Integrating \eqref{eq:ShallowWater} along the flow ($y$-coordinate) results in the Bernoulli function describing the sum of kinetic and potential energy of the system which is conserved for an isolated single layer except for dissipative regions like hydraulic jumps:
\begin{equation}
    B = \frac{v^2}{2} + g^\prime \delta + g^\prime h\ .
\end{equation}
The change in the energy flux associated with the transport of the Bernoulli function $F=Q B$, or Bernoulli flux, between upstream and downstream of a dissipative region over a flat bottom is
\begin{align}
\begin{split}
\Delta F &= Q_u B_u -Q_d B_d = Q(B_u - B_d) \\
    &= v_u d_u (\frac{v_u^2}{2} + g^\prime \delta_u - \frac{v_d^2}{2} -g^\prime \delta_d)\ ,
\end{split}
\label{eq:BernoulliNoEntrainment}
\end{align}
with subscripts $u$ and $d$ denoting upstream and downstream of a jump.
If entrainment is allowed then the volume flux changes and the drop is
\begin{equation}
\Delta F = v_u \delta_u (\frac{v_u^2}{2} + g^\prime_u \delta_u) - v_d \delta_d(\frac{v_d^2}{2} -g^\prime_d \delta_d)\ .
\label{eq:BernoulliEntrainment}
\end{equation}
Note that in \eqref{eq:BernoulliEntrainment} the bottom depth is the same between upstream and downstream.
We can express the energy drop including changes in bottom depth by adding the $h$ term:
\begin{equation}
\Delta F = \frac{v_u^3 \delta_u}{2} + v_u g^\prime_u (\delta_u^2 + h_u \delta_u) - \frac{v_d^3 \delta_d}{2} -v_d g^\prime_d (\delta_d^2 + h_d \delta_d)\ .
\label{eq:BernoulliEntrainmentVariableBottom}
\end{equation}
We can calculate $\Delta F$ following \eqref{eq:BernoulliEntrainmentVariableBottom} for various points upstream and downstream in model and observations, however, as we define a single layer $g^\prime$ and single layer velocity $v$, the results will be somewhat coarse.
As we will show in section~\ref{sec:Results}\ref{ssec:EnergyBudgetResults}, the drop in Bernoulli flux only converges to a meaningful result when sufficiently averaged in time.
% Furthermore, in this Bernoulli formulation of the overflow energetics we lack any information on how the energy loss occurs; whether it all dissipates within the overflow layer or whether some energy may radiate upwards in the form of an internal gravity wave energy flux.
We thus turn to a more detailed description of the overflow energetics in the next section.
Nevertheless, we expect results from these two approaches to be broadly comparable with each other.

@@

Vertical pressure work and internal wave fluxes, the small-scale subset of the pressure work term, are mostly upward and concentrated within the dense bottom layer.
The total pressure work term integrated along isopycnals between kilometer~0 and 17 differs quite substantially between observations and model (Fig.~\ref{fig:VerticalIWIntegrated}a), especially for the densest layers where it shows a downward flux of energy for the 2012 towyo. 
Fluxes within the dense layer vary between -4 and 4\,kW\,m$^{-1}$. 
Both towyo sections and the model show upward energy flux due to the pressure work term near the interface and diminishing magnitudes towards and beyond the interface.
The disagreement between the two towyos and the model makes the vertical pressure work term the least consistent term in the energy budget.
Integrated along isopycnals, the vertical energy flux due to smaller-scale internal waves is directed upwards and reaches between 1~and 2\,kW\,m$^{-1}$ within the bottom layer in the observations and somewhat smaller magnitudes in the model (Fig.~\ref{fig:VerticalIWIntegrated}b).
In both cases, vertical internal wave fluxes diminish close to zero past the upper interface of the overflow layer, indicating that the high frequency waves do not radiate much energy aloft outside the overflow layer.

@@

\begin{figure*}
\centerline{\includegraphics[width=27pc]{bottom_pressure.pdf}}
\caption{Pressure anomaly in observations and model.
a) Baroclinic pressure anomaly (colors) and isopycnals (contours) in the 2012 towyo transect.
The thick contour shows the $\sigma_4=45.94$\,kg\,m$^{-3}$ isopycnal previously defined as the upper layer interface.
The shaded area above the bottom shows depths not reached by the CTD observations and where constant density was assumed in the bottom pressure calculation.
b) Baroclinic bottom pressure in observations and model offset by constant factors for visualization purposes.
Faint black lines show bottom pressure for each model time step of the analysis period.
The pink line shows the time mean model bottom pressure for the same period.
Blue and purple lines show bottom pressure in the observations. 
}
\label{fig:BottomPressure}
\end{figure*}

@@

\appendix[B]
\appendixtitle{Small-scale internal wave fluxes}
We work with locally defined velocity and pressure perturbations $\mathbf{u''}$ and $p''$ as detailed in section~\ref{sec:Energetics}\ref{sec:EnergyEquation} in \eqref{eq:BaroclinicPressureDecomposition} and \eqref{eq:BaroclinicoVelocityDecomposition} to investigate the role of smaller-scale internal waves for the energetics of the dense layer.
Varying the window size in the calculation of local mean profiles between 3 and 8\,km in the model changes the magnitude of the horizontally integrated vertical wave flux as shown in Fig.~\ref{fig:VerticalIWIntegrated}b by about a factor of two with diminishing energy flux for smaller window sizes corresponding to smaller lateral scales.

>>>>

\abstract{
The energy and momentum balance of an abyssal overflow across a major sill in the Samoan Passage is estimated from two highly resolved towed sections, set 16 months apart, and results from a two-dimensional numerical simulation.
Driven by the density anomaly across the sill, the flow is relatively steady.
The system gains energy from divergence of horizontal pressure work $\mathcal{O}(5)$\,kW\,m$^{-1}$ and flux of available potential energy $\mathcal{O}(2)$\,kW\,m$^{-1}$.
Approximately half of these gains are transferred into kinetic energy while the other half is lost to turbulent dissipation, bottom drag, and divergence in vertical pressure work.
Small-scale internal waves emanating downstream of the sill within the overflow layer radiate $\mathcal{O}(1)$\,kW\,m$^{-1}$ upwards but dissipate most of their energy within the dense overflow layer and at its upper interface.
The strongly sheared and highly stratified upper interface acts as a critical layer inhibiting any appreciable upward radiation of energy via topographically generated lee waves.
Form drag of $\mathcal{O}(2)$\,N\,m$^{-2}$, estimated from the pressure drop across the sill, is consistent with energy lost to dissipation and internal wave fluxes.
The topographic drag removes momentum from the mean flow, slowing it down and feeding a counter current aloft.
The processes discussed in this study combine to convert about one third of the energy released from the cross-sill density difference into turbulent mixing within the overflow and at its upper interface.
The observed and modeled vertical momentum flux divergence sustains gradients in shear and stratification, thereby maintaining an efficient route for abyssal water mass transformation downstream of this Samoan Passage sill.
}

@@

In this study we estimate the energy and momentum budget of flow across a major sill in the Samoan Passage using high horizontal resolution, towed, observations.
Results from a two-dimensional model are used to corroborate the analysis.
In the following, we give a short overview of the abyssal flow through the Samoan Passage and one of its major overflows (section~\ref{sec:ExperimentalDetails}\ref{sec:StudyRegion}), present towed observations of this overflow (section~\ref{sec:ExperimentalDetails}\ref{sec:Observations}),  and outline the setup of a two-dimensional numerical model simulating the dense overflow (section~\ref{sec:ExperimentalDetails}\ref{sec:Model}) to help interpret the observations.
After discussing the energy of the flow in terms of the Bernoulli equation (section~\ref{sec:Energetics}\ref{sec:Bernoulli}), a baroclinic energy equation is introduced (section~\ref{sec:Energetics}\ref{sec:EnergyEquation}).
Flow steadiness is discussed as a precursor to the following analysis in section~\ref{sec:Results}\ref{ssec:FlowSteadiness}.
Both energy frameworks are applied to the observed and modeled overflow to obtain energy budgets in section~\ref{sec:Results}\ref{ssec:EnergyBudgetResults}.
Form drag is calculated and evaluated against the energy budgets (\ref{sec:Results}\ref{ssec:FormDrag}).
Upward momentum flux estimates are presented in section~\ref{sec:Results}\ref{ssec:MomentumFluxes}.
The results are discussed and compared to observations of other high drag flows in section~\ref{sec:Discussion}.

@@

\begin{figure*}
\centerline{\includegraphics[width=39pc]{model_snapshot.pdf}}
\caption{Model snapshot after 100 hours spinup time. Potential temperature is contoured in all four panels in 0.05$^{\circ}$C intervals starting at 0.7$^{\circ}$C. Thicker contours show the 0.8 and 0.9$^{\circ}$C isotherms.
(a) Potential temperature.
(b) Horizontal velocity.
(c) Vertical velocity.
(d) Turbulent dissipation $\varepsilon$ based on a parameterization acting on vertical density instabilities \citep[KL10,][]{klymaklegg10}.
}
\label{fig:ModelSnapshot}
\end{figure*}

@@

\subsection{Bernoulli flux}\label{sec:Bernoulli}
Treating the overflow as a single layer flow with the waters above at rest, we start out with the steady shallow water equations in one dimension:
\begin{equation}
    v \frac{\partial v}{\partial y} + g^\prime \frac{\partial}{\partial y} (\delta + h) = 0\ ,
    \label{eq:ShallowWater}
\end{equation}
where $\delta$ is the thickness of the layer, $v$ is the horizontal layer velocity along coordinate $y$, $h$ is the elevation of the topography, and $g^\prime=g \Delta \rho / \rho$ expresses the density difference $\Delta \rho$ between the bottom layer and waters aloft.
Neglecting entrainment leads to constant volume transport $Q$ of the dense bottom layer:
\begin{equation}
\frac{\partial (v \delta)}{\partial y} = \frac{\partial Q}{\partial y} = 0\ .
\end{equation}
Integrating \eqref{eq:ShallowWater} along the flow ($y$-coordinate) results in the Bernoulli function describing the sum of kinetic and potential energy of the system which is conserved for an isolated single layer except for dissipative regions like hydraulic jumps:
\begin{equation}
    B = \frac{v^2}{2} + g^\prime \delta + g^\prime h\ .
\end{equation}
The change in the energy flux associated with the transport of the Bernoulli function $F=Q B$, or Bernoulli flux, between upstream and downstream of a dissipative region over a flat bottom is
\begin{align}
\begin{split}
\Delta F &= Q_u B_u -Q_d B_d = Q(B_u - B_d) \\
    &= v_u \delta_u (\frac{v_u^2}{2} + g^\prime \delta_u - \frac{v_d^2}{2} -g^\prime \delta_d)\ ,
\end{split}
\label{eq:BernoulliNoEntrainment}
\end{align}
with subscripts $u$ and $d$ denoting upstream and downstream of a jump.
If entrainment is allowed then the volume flux changes and the drop is
\begin{equation}
\Delta F = v_u \delta_u (\frac{v_u^2}{2} + g^\prime_u \delta_u) - v_d \delta_d(\frac{v_d^2}{2} -g^\prime_d \delta_d)\ .
\label{eq:BernoulliEntrainment}
\end{equation}
Note that in \eqref{eq:BernoulliEntrainment} the bottom depth is the same between upstream and downstream.
We can express the energy drop including changes in bottom depth by adding the $h$ term:
\begin{equation}
\Delta F = \frac{v_u^3 \delta_u}{2} + v_u g^\prime_u (\delta_u^2 + h_u \delta_u) - \frac{v_d^3 \delta_d}{2} -v_d g^\prime_d (\delta_d^2 + h_d \delta_d)\ .
\label{eq:BernoulliEntrainmentVariableBottom}
\end{equation}
We can calculate $\Delta F$ following \eqref{eq:BernoulliEntrainmentVariableBottom} for various points upstream and downstream in model and observations, however, as we define a single layer $g^\prime$ and single layer velocity $v$, the results will be somewhat coarse.
As we will show in section~\ref{sec:Results}\ref{ssec:EnergyBudgetResults}, the drop in Bernoulli flux only converges to a meaningful result when sufficiently averaged in time.
% Furthermore, in this Bernoulli formulation of the overflow energetics we lack any information on how the energy loss occurs; whether it all dissipates within the overflow layer or whether some energy may radiate upwards in the form of an internal gravity wave energy flux.
We thus turn to a more detailed description of the overflow energetics in the next section.
Nevertheless, we expect results from these two approaches to be broadly comparable with each other.

@@

Vertical pressure work and internal wave fluxes, the small-scale subset of the pressure work term, are mostly upward and concentrated within the dense bottom layer.
The total pressure work term integrated along isopycnals between kilometer~0 and 17 differs quite substantially between observations and model (Fig.~\ref{fig:VerticalIWIntegrated}a), especially for the densest layers where it shows a downward flux of energy for the 2012 towyo. 
Pressure work within the dense layer varies between -4 and 4\,kW\,m$^{-1}$. 
Both towyo sections and the model show upward energy flux due to the pressure work term near the interface and diminishing magnitudes towards and beyond the interface.
The disagreement between the two towyos and the model makes the vertical pressure work term the least consistent term in the energy budget.
Integrated along isopycnals, the vertical energy flux due to smaller-scale internal waves is directed upwards and reaches between 1~and 2\,kW\,m$^{-1}$ within the bottom layer in the observations and somewhat smaller magnitudes in the model (Fig.~\ref{fig:VerticalIWIntegrated}b).
In both cases, vertical internal wave fluxes diminish close to zero past the upper interface of the overflow layer, indicating that the high frequency waves do not radiate much energy aloft outside the overflow layer.

@@

\begin{figure*}
\centerline{\includegraphics[width=27pc]{bottom_pressure.pdf}}
\caption{Pressure anomaly in observations and model.
a) Baroclinic pressure anomaly (colors) and isopycnals (contours) in the 2012 towyo transect.
The thick contour shows the $\sigma_4=45.94$\,kg\,m$^{-3}$ isopycnal previously defined as the upper layer interface.
The shaded area above the bottom shows depths not reached by the CTD observations and where constant density was assumed in the bottom pressure calculation.
b) Baroclinic bottom pressure in observations and model offset by constant factors (model:~60\,N\,m$^{-2}$, towyo~2012:~-130\,N\,m$^{-2}$, towyo~2014:~-150\,N\,m$^{-2}$) for visualization purposes.
This relatively large offset in absolute baroclinic bottom pressure between model and observations is due to model initialization with the observed density structure and partial draining of the initial stratification during model spinup as discussed in section~\ref{sec:ExperimentalDetails}\ref{sec:Model}.
Faint black lines show bottom pressure for each model time step of the analysis period.
The pink line shows the time mean model bottom pressure for the same period.
Blue and purple lines show bottom pressure in the observations. 
}
\label{fig:BottomPressure}
\end{figure*}

@@

\appendix[B]
\appendixtitle{Small-scale internal wave fluxes}
We work with locally defined velocity and pressure perturbations $\mathbf{u''}$ and $p''$ as detailed in section~\ref{sec:Energetics}\ref{sec:EnergyEquation} in \eqref{eq:BaroclinicPressureDecomposition} and \eqref{eq:BaroclinicoVelocityDecomposition} to investigate the role of smaller-scale internal waves for the energetics of the dense layer.
Varying the window size in the calculation of local mean profiles between 3 and 8\,km in the model changes the magnitude of the horizontally integrated vertical wave flux by about a factor of two with diminishing energy flux for smaller window sizes corresponding to smaller lateral scales.
